Optimal Finite Element Integration

### 2015-07-29 TOMS_Paper/assembly.tex
Minor fixes

\subsection{Operation-minimum synthesis}
Definition~\ref{def:optimality} states that a necessary condition for a loop nest synthesis to be optimal is that the number of operations in the innermost loop is minimum. We now discuss how we can systematically achieve this.

Eliminating sharing from the multilinear loops does not suffice to reach nest optimality. In fact, we wonder whether the reduction imposed by $L_i$ could be pre-evaluated outside of $N$, as already suggested in~\ref{def:optimality}, thus potentially reducing the operation count.

To answer this question, we make use of a result -- the foundation of tensor contraction mode -- in~\cite{Kirby:TC}. Essentially, multilinear forms can be seen as sums of monomials, each monomial being an integral over the equation domain of products (of derivatives) of functions from discrete spaces; such monomials can always be reduced to a product of two tensors (see Section~\ref{sec:background}). We interpret this result at the loop nest level: with the input as in Figure~\ref{code:loopnest}, we can always dissect \texttt{F} into distinct sub-expressions (the monomials). Each sub-expression is then factorized so as to split constant from $[L_i, L_j, L_k]$-dependent terms, the latter ones are hoisted outside of $N$ and finally pre-evaluated. As part of this pre-evaluation, the reduction induced by $L_i$ vanishes. In the following, we simply refer to this special sort of code hoisting as "\textit{pre-evaluation}".

@@

If we estimate that, for a given monomial, pre-evaluation will decrease the workload, then the corresponding sub-expression is extracted, a sequence of transformation steps -- involving expansion, factorization, code motion -- takes place (details in Section~\ref{sec:codegen}, and the evaluation eventually performed. The result is a set of $n$-dimensional tables (these can be seen as "slices" of the reference tensor at the math level), $n$ being the arity of the multilinear form. Identical tables are mapped to the same temporary. Eventually, sharing is removed in all resulting expressions by applying a procedure as described in Proposition~\ref{prop:optimal-mln}.

>>>>

\subsection{Operation Minimization under Memory Constraints}
Definition~\ref{def:optimality} states that a necessary condition for a loop nest synthesis to be optimal is that the number of operations in all innermost loops is minimum. We now discuss how we can systematically achieve this.

Eliminating sharing from the multilinear loops does not suffice for nest optimality. In fact, we wonder whether the reduction imposed by $L_i$ could be pre-evaluated outside of $N$, as already suggested in~\ref{def:optimality}, thus potentially reducing the operation count.

To answer this question, we make use of a result -- the foundation of tensor contraction mode -- from~\citeN{Kirby:TC}. Essentially, multilinear forms can be seen as sums of monomials, each monomial being an integral over the equation domain of products (of derivatives) of functions from discrete spaces; such monomials can always be reduced to a product of two tensors (see Section~\ref{sec:background}). We interpret this result at the loop nest level: with the input as in Figure~\ref{code:loopnest}, we can always dissect \texttt{F} into distinct sub-expressions (the monomials). Each sub-expression is then factorized so as to split constant from $[L_i, L_j, L_k]$-dependent terms, the latter ones are hoisted outside of $N$, and finally pre-evaluated. As part of this pre-evaluation, the reduction induced by $L_i$ vanishes. In the following, we simply refer to this special sort of code hoisting as "\textit{pre-evaluation}".

@@

If we estimate that, for a given monomial, pre-evaluation will decrease the workload, then the corresponding sub-expression is extracted, a sequence of transformation steps -- involving expansion, factorization, code motion -- takes place (details in Section~\ref{sec:codegen}, and the evaluation eventually performed. The result is a set of $n$-dimensional tables (these can be seen as "slices" of the reference tensor at the math level), $n$ being the arity of the multilinear form. Identical tables are mapped to the same temporary. Eventually, sharing is removed from all resulting expressions by applying a procedure as described in Proposition~\ref{prop:optimal-mln}.
